---
title: "DVProject1"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# options(warn=-1)
library(tidyverse)
library(ggplot2)
library(scales)
library(cowplot)
library(wesanderson)

# datapath <- '/Users/viv/Downloads/DATA'
```  

```{r, include=FALSE}
df.students <- read_csv('StudentsPerformance.csv')
```




## Part 3

```{r}
plot3 <- ggplot(data = df.students, aes(x=`reading score`, y=100*(..count../sum(..count..)))) +
  geom_histogram(color="black", fill="paleturquoise", boundary= 0, binwidth = 5, closed="left")  +
  scale_y_continuous(expand=c(0,0)) + coord_cartesian(ylim = c(0, 15)) +
  labs(title="Reading Scores", subtitle = "Based on 1,000 US High School Students", x="Score (%)", y="Percent (%)", caption = "Data From Kaggle: Student Performance")  +
  theme_bw() +
  theme(plot.title=element_text(hjust=0.5), plot.subtitle=element_text(hjust=0.5))

plot3 + geom_vline(aes(xintercept = mean(`reading score`)), color = "black", linetype = "dashed", size = 0.8)
```

The distribution of reading scores is skewed left. The reading scores are centered around 69% with a mean of 69% and median of 70%. The reading scores range from 17% to 100% and there are no outliers.



## Part 4

```{r}
plot4 <- ggplot(data=df.students, aes(x=`math score`, y = 100*(..count../sum(..count..)), fill=`test preparation course`)) +
  geom_histogram(color = 'black', binwidth = 5) + labs(title="Math Scores by Test Preparation", subtitle = "Based on 1,000 US High School Students", x="Score (%)", y="Percent (%)", caption = "Data From Kaggle: Student Performance")  + 
  theme_bw() + theme(plot.title=element_text(hjust=0.5), plot.subtitle=element_text(hjust=0.5)) +
  facet_wrap(vars(`test preparation course`)) + guides(fill='none') 

plot4 + scale_fill_manual(values = c('#9EB9A5', '#D27575')) + geom_vline(aes(xintercept = mean(`math score`)), color = "black", linetype = "dashed", size = 0.5)
```

Both of the distributions of math scores are skewed left. The math scores, regardless of test preparation, centered around 66%. The math scores of students that completed test preparation centered around 69%, whereas the math scores of students that did not complete test preparation centered around 64%. The distributions also indicate that less students completed test preparation (358 students) than students that did not complete test preparation (642 students). The math scores of students that completed test preparation ranges from 23% to 100% and the math scores of students that did not complete test preparation ranges from 0% to 100%. There are no outliers.
